Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-01-30 numerical-conduction.tex
First draft of thermal conduction section done.

We've got both kinds of conduction: isotropic AND anisotropic.  How
great is that?

@@

where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
/ |\nabla T|$ is the characteristic length scale of the local
temperature gradient.

It is important to note that the conductivity can be strongly affected
by the presence of magnetic fields, which may strongly suppress heat
flow.  The strength and confuction of magnetic fields in the
intracluster medium is poorly understood, and it is unclear to what
extent these magnetic fields will suppress conduction.  For the
purposes of the work described in this paper, we assume that the
conductivity of the plasma can be described in terms of an effective
conductivity, which can be expressed as a fraction f$_{Sp}$ of the
Spitzer conductivity (where f$_{Sp} \leq 1.0$ are considered physical
values).

>>>>

@@

where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
/ |\nabla T|$ is the characteristic length scale of the local
temperature gradient.  Furthermore, it is important to note that the
overall level of thermal conductivity can be strongly affected
by the presence of magnetic fields, which may strongly suppress heat
flow.  The strength and confuction of magnetic fields in the
intracluster medium is poorly understood, and it is unclear to what
extent these magnetic fields will suppress conduction.  For the
purposes of the work described in this paper, we assume that the
conductivity of the plasma can be described in terms of an effective
conductivity, which can be expressed as a fraction f$_{Sp}$ of the
Spitzer conductivity (where f$_{Sp} \leq 1.0$ are considered physical
values).

Thermal conduction in a
plasma can be strongly affected by the presence of magnetic field
lines, which may strongly suppress heat flow perpendicular to the
magnetic field.  In that case, we allow for heat transport only
parallel to
the magnetic field lines in magnetohydrodynamic simulations.
Mathematically, this is treated as:

\begin{equation}
\mathbb{F} = -\kappa_{sp} \^{b} (\^{b} \cdot \nabla T)
\end{equation}

Where $\^{b}$ is the unit vector denoting the direction of the
magnetic field.  As with the isotropic thermal conduction, we allow a
multiplicative factor f$_{Sp}$ to take into account the possible
behavior of magnetic fields below the resolution limit of the
simulation.

Both isotropic and anisotropic thermal conduction in Enzo are treated
in an operator-split manner.  Furthermore, within the heat transport
module, transport along the x, y, and z directions are computed in a
directionally-split fashion, with heat flux along each direction
calculated at the + and - faces of the cell using the arithmetic mean
of the cell-centered temperature in cells $n$ and $n+1$ or $n-1$ and
$n$, respectively (empirically, this is more stable than taking the
geometric mean of the cell-centered temperatures).  The addition of
transport along magnetic field lines requires the calculation of
cross-terms in the temperature derivatives at cell faces, which can
result in spurious oscillations in the temperature field in regions
where the temperature gradient is strong in more than one spatial
direction.  Controlling these oscillations requires the addition of a
flux limiter for calculations of the temperature field.  In this case,
we choose the monotonized central difference flux limiter \citep{1977JCoPh..23..263V}, which serves to maintain numerical stability without
sacrificing substantial speed or accuracy.
